---
title: "PC219 Crosslinking MS Exam Question"
author: "Mike Trnka"
date: "`r format(Sys.time(), '%b %d %Y')`"
output:
   html_document:
      toc: TRUE
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
# published at:
# https://mtrnka.github.io/pc219exam/CLMS_exam.html
# Spectrum is on MS-Viewer
# code: si9qwtwl9e
# https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?report_title=MS-Viewer&search_key=si9qwtwl9e&search_name=msviewer

knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(flextable)

pep.assignments <- tibble(
   assignment = c("#1","#2","#3","#4"),
   url = c(
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=AVDIPHMDIEALK%28%2bDSSO%2a%29K&s=1&sequence2=K%28%2bDSSO%2a%29PRHQEGEIFDTEK&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&",
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=LITEDVQGK%28%2bDSSO%2a%29NC%28Carbamidomethyl%29LTNFHGMDLTR&s=1&sequence2=MLAIK%28%2bDSSO%2a%29R&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&",
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=LITEDVQGK%28%2bDSSO%2a%29NC%28Carbamidomethyl%29LTNFHGMDLTR&s=1&sequence2=TK%28%2bDSSO%2a%29LQNK&s2=1&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&",
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=LITEDVQGK%28%2BDSSO*%29NC%28Carbamidomethyl%29LTNFHGMDLTR&s=1&sequence2=ELATK%28%2BDSSO*%29LG&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&"
   ),
   peptide.1 = c("AVDIPHMDIEALKK","LITEDVQGKNCLTNFHGMDLTR","LITEDVQGKNCLTNFHGMDLTR",
                 "LITEDVQGKNCLTNFHGMDLTR"),
   peptide.2 = c("KPRHQEGEIFDTEK","MLAIKR","TKLQNK","ELATKLG"),
   protein.1 = c("uL1","eS1","eS1","eS1"),
   prot.1.url = c("https://www.uniprot.org/uniprotkb/G1TY53/entry",
                  "https://www.uniprot.org/uniprotkb/G1SS70/entry",
                  "https://www.uniprot.org/uniprotkb/G1SS70/entry",
                  "https://www.uniprot.org/uniprotkb/G1SS70/entry"),
   protein.2 = c("eL6","eL42","uL16","eL8"),
   prot.2.url = c("https://www.uniprot.org/uniprotkb/G1SKF7/entry",
                  "https://www.uniprot.org/uniprotkb/G1U344/entry",
                  "https://www.uniprot.org/uniprotkb/B7NZQ2/entry",
                  "https://www.uniprot.org/uniprotkb/G1STW0/entry"),
   residue.1 = c(91,94,94,94),
   residue.2 = c(227,86,141,317),
   chainID.1 = c("w3","C2","C2","C2"),
   chainID.2 = c("E3","o3","I3","G3"),
   module.1 = c("60S Ribosome","40S Ribosome","40S Ribosome","40S Ribosome"),
   module.2 = c("60S Ribosome","60S Ribosome","60S Ribosome","60S Ribosome")
)
pep.flex <- flextable(pep.assignments, col_keys=c(names(pep.assignments)[c(-2,-6,-8)]))
pep.flex <- compose(x = pep.flex, j = "assignment", 
                    value = as_paragraph(hyperlink_text(x = assignment, url = url))
)
pep.flex <- compose(x = pep.flex, j = "protein.1", 
                    value = as_paragraph(hyperlink_text(x = protein.1, url = prot.1.url))
)
pep.flex <- compose(x = pep.flex, j = "protein.2", 
                    value = as_paragraph(hyperlink_text(x = protein.2, url = prot.2.url))
)

```

For this exam question we will look at some CLMS data that was obtained by
crosslinking 80S Ribosome from Rabbit with the cleavable crosslinker
[DSSO](https://www.thermofisher.com/order/catalog/product/A33545){target="_blank"}.

We are interested in one particular MS2 HCD product ion spectrum from
this experiment. We have identified four possible crosslinked sequences
that match the precursor ion (m/z = **1150.8955^3+^**) and the product
ions. Each of the four assignments links to an annotated product ion
spectrum showing the ions that support that assignment. By clicking on
the link in the "assignment" column, you will be able to see detailed
information about the match between the spectrum and the sequences. Some
useful things to consider are the calculated mono-isotopic MH of the
precursor ion, the mass error of the product ion matches, how much of
the sequence of each peptide is matched by product ions, how much noise
(unmatched ions) are in the annotation, etc.

DSSO is a cleavable crosslinker, which can cleave on either side of the
sulfoxide. The ions which contain a DSSO can each produce either a thiol
or an alkene modification. In Protein Prospector these are annotated
with the symbols \* and #.

```{r table, echo=F}
big_border = fp_border_default(color="grey85", width = 2)
small_border = fp_border_default(color="gray", width = 1)

pep.flex <- border_inner_h(pep.flex, part="all", border = small_border )
pep.flex <- border_inner_v(pep.flex, part="all", border = small_border )
pep.flex <- border_outer(pep.flex, part="all", border = big_border )
pep.flex %>%
   bold(part = "header") %>%
   color(part = "body", j = c("assignment", "protein.1", "protein.2"), color = "dodgerblue")
```

### Question 1:

When the Orbitrap analyzer is callibrated well, it typically has a mass
accuracy of less than 5 ppm. Compare the calculated MH^+^ of the
putative crosslink matches to the experimental precursor ion. Are any of
the matches within this precursor mass tolerance.

### Question 2

Do any of the putative assignments match the diagnostic DSSO ion pairs?
The DSSO paper might be helpful [@kao2011b].

### Question 3

Look at the spectral annotation and also look underneath in the "Main
Sequence Ions" area to see which parts of the peptide sequences are
matched by **y-** and **b**-ions.

Which of the putative crosslinks is most likely to be assigned
correctly?

### Question 4

There is a reference structure for the rabbit Ribosome, PDB code:
[6hcj](https://www.rcsb.org/structure/6HCJ){target="_blank"}

The table contains the residue numbers and the PDB chain codes for
putative crosslinks. Measure the C$\alpha$ to C$\alpha$ distances of the
crosslink assignments. Note: In
[ChimeraX](https://www.cgl.ucsf.edu/chimerax/){target="_blank"}, you can load the pdb
file and use the distance command in the command line like so:

    distance /w3:91@CA /E3:224@CA

For assignment #1, residue 227 on Ribosome Protein eL6 is not resolved in
the pdb structure so use residue 224, the closest mapped residue, as I
did in the example above.

Are any of the crosslink assignments consistent with the expected
distance of the DSSO crosslinker (\< 30 Å)?

### Question 5

What aspects of the spectral annotation were most useful in determining
whether or not the crosslink was assigned correctly?


### References: