---
title: "PC219 Crosslinking MS Exam Question Answer Key"
author: "Mike Trnka"
date: "`r format(Sys.time(), '%b %d %Y')`"
output:
   html_document:
      toc: TRUE
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---
```{r setup, include=FALSE}
# published at:
# https://mtrnka.github.io/pc219exam/CLMS_exam.html
# Spectrum is on MS-Viewer
# code: si9qwtwl9e
# https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?report_title=MS-Viewer&search_key=si9qwtwl9e&search_name=msviewer

knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
library(flextable)

pep.assignments <- tibble(
   assignment = c("#1","#2","#3","#4"),
   url = c(
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=AVDIPHMDIEALK%28%2bDSSO%2a%29K&s=1&sequence2=K%28%2bDSSO%2a%29PRHQEGEIFDTEK&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&",
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=LITEDVQGK%28%2bDSSO%2a%29NC%28Carbamidomethyl%29LTNFHGMDLTR&s=1&sequence2=MLAIK%28%2bDSSO%2a%29R&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&",
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=LITEDVQGK%28%2bDSSO%2a%29NC%28Carbamidomethyl%29LTNFHGMDLTR&s=1&sequence2=TK%28%2bDSSO%2a%29LQNK&s2=1&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&",
   "https://msviewer.ucsf.edu/cgi-bin/mssearch.cgi?search_name=msproduct&output_type=HTML&report_title=MS-Product&version=6.5.0%20Basic&data_source=Data%20From%20File&data_filename=%2Fvar%2Flib%2Fprospector%2Fweb%2Fresults%2Fmsviewer%2Fs%2Fi%2Fsi9qwtwl9e%2FDSSOstar_rRibo%2FZ20200519-53_FTMSms2hcd.mgf&use_instrument_ion_types=1&msms_min_precursor_mass=0&instrument_name=ESI-Q-high-res&display_graph=1&msms_parent_mass_tolerance=200&msms_parent_mass_tolerance_units=ppm&fragment_masses_tolerance=30&fragment_masses_tolerance_units=ppm&msms_pk_filter=Max%20MSMS%20Pks&msms_max_peaks=100&scan_number=19752&max_charge=3&msms_precursor_charge=3&sequence=LITEDVQGK%28%2BDSSO*%29NC%28Carbamidomethyl%29LTNFHGMDLTR&s=1&sequence2=ELATK%28%2BDSSO*%29LG&s2=1&count_pos_z=Count%20Basic%20AA&link_search_type=DSSO*&"
   ),
   peptide.1 = c("AVDIPHMDIEALKK","LITEDVQGKNCLTNFHGMDLTR","LITEDVQGKNCLTNFHGMDLTR",
                 "LITEDVQGKNCLTNFHGMDLTR"),
   peptide.2 = c("KPRHQEGEIFDTEK","MLAIKR","TKLQNK","ELATKLG"),
   protein.1 = c("uL1","eS1","eS1","eS1"),
   prot.1.url = c("https://www.uniprot.org/uniprotkb/G1TY53/entry",
                  "https://www.uniprot.org/uniprotkb/G1SS70/entry",
                  "https://www.uniprot.org/uniprotkb/G1SS70/entry",
                  "https://www.uniprot.org/uniprotkb/G1SS70/entry"),
   protein.2 = c("eL6","eL42","uL16","eL8"),
   prot.2.url = c("https://www.uniprot.org/uniprotkb/G1SKF7/entry",
                  "https://www.uniprot.org/uniprotkb/G1U344/entry",
                  "https://www.uniprot.org/uniprotkb/B7NZQ2/entry",
                  "https://www.uniprot.org/uniprotkb/G1STW0/entry"),
   residue.1 = c(91,94,94,94),
   residue.2 = c(227,86,141,317),
   chainID.1 = c("w3","C2","C2","C2"),
   chainID.2 = c("E3","o3","I3","G3"),
   module.1 = c("60S Ribosome","40S Ribosome","40S Ribosome","40S Ribosome"),
   module.2 = c("60S Ribosome","60S Ribosome","60S Ribosome","60S Ribosome")
)
pep.flex <- flextable(pep.assignments, col_keys=c(names(pep.assignments)[c(-2,-6,-8)]))
pep.flex <- compose(x = pep.flex, j = "assignment", 
                    value = as_paragraph(hyperlink_text(x = assignment, url = url))
)
pep.flex <- compose(x = pep.flex, j = "protein.1", 
                    value = as_paragraph(hyperlink_text(x = protein.1, url = prot.1.url))
)
pep.flex <- compose(x = pep.flex, j = "protein.2", 
                    value = as_paragraph(hyperlink_text(x = protein.2, url = prot.2.url))
)

```

For this exam question we will look at some CLMS data that was obtained by
crosslinking 80S Ribosome from Rabbit with the cleavable crosslinker
[DSSO](https://www.thermofisher.com/order/catalog/product/A33545){target="_blank"}.

We are interested in one particular MS2 HCD product ion spectrum from
this experiment. We have identified four possible crosslinked sequences
that match the precursor ion (m/z = **1150.8955^3+^**) and the product
ions. Each of the four assignments links to an annotated product ion
spectrum showing the ions that support that assignment. By clicking on
the link in the "assignment" column, you will be able to see detailed
information about the match between the spectrum and the sequences. Some
useful things to consider are the calculated mono-isotopic MH of the
precursor ion, the mass error of the product ion matches, how much of
the sequence of each peptide is matched by product ions, how much noise
(unmatched ions) are in the annotation, etc.

DSSO is a cleavable crosslinker, which can cleave on either side of the
sulfoxide. The ions which contain a DSSO can each produce either a thiol
or an alkene modification. In Protein Prospector these are annotated
with the symbols \* and #.

```{r table, echo=F}
big_border = fp_border_default(color="grey85", width = 2)
small_border = fp_border_default(color="gray", width = 1)

pep.flex <- border_inner_h(pep.flex, part="all", border = small_border )
pep.flex <- border_inner_v(pep.flex, part="all", border = small_border )
pep.flex <- border_outer(pep.flex, part="all", border = big_border )
pep.flex %>%
   bold(part = "header") %>%
   color(part = "body", j = c("assignment", "protein.1", "protein.2"), color = "dodgerblue")
```

### Question 10:

When the Orbitrap analyzer is callibrated well, it typically has a mass
accuracy of less than 5 ppm. Compare the calculated MH^+^ of the
putative crosslink matches to the experimental precursor ion. Are any of
the matches within this precursor mass tolerance.

[**Answer**]{style="color:red"}

[The MS-Product links for the assignment include the calculated
monoisotopic MH^3+^ values (M~c~). The experimentally measured m/z was
given above (M~e~). Mass accuracy can be calculated from: 1e^6^ x
(M~e~ - M~c~) / M~c~]{style="color:red"}

```{r ppm, echo=F}
pep.ppm <- pep.assignments %>% select(1:4)
pep.ppm$mh3.experimental = 1150.8955
pep.ppm$mh3.calculated = c(1150.9059, 1150.9049, 1150.8987, 1150.8949)
pep.ppm <- pep.ppm %>%
   mutate(ppm = map2_dbl(mh3.experimental, mh3.calculated, 
                         function(x,y) {round(1e6 * (x - y) / y, 2)}))
pep.ppm.flex <- flextable(pep.ppm, col_keys=c(names(pep.ppm)[c(-2)]))
pep.ppm.flex <- compose(x = pep.ppm.flex, j = "assignment", 
                    value = as_paragraph(hyperlink_text(x = assignment, url = url))
)
pep.ppm.flex <- border_inner_h(pep.ppm.flex, part="all", border = small_border )
pep.ppm.flex <- border_inner_v(pep.ppm.flex, part="all", border = small_border )
pep.ppm.flex <- border_outer(pep.ppm.flex, part="all", border = big_border )
pep.ppm.flex %>%
   bold(part = "header") %>%
   color(part = "body", j = c("assignment"), color = "dodgerblue")

```

[Only assignments #3 and #4 are within the 5 ppm mass acuracy tolerance]{style="color:red"}

### Question 11

Do any of the putative assignments match the diagnostic DSSO ion pairs?
The DSSO paper might be helpful [@kao2011b].

[**Answer**]{style="color:red"}

[Assignments #2, #3, #4 have all 4 diagnostic DSSO ions matched. These are the MH* and MH# ions in the annotations]{style="color:red"}

### Question 12

Look at the spectral annotation and also look underneath in the "Main
Sequence Ions" area to see which parts of the peptide sequences are
matched by **y-** and **b**-ions.

Which of the putative crosslinks is most likely to be assigned
correctly?

[**Answer**]{style="color:red"}

[Assignments #1 only matches a few random product ions from each sequnce. 
Assignments #2, #3, and #4 all have nearly complete production ion coverage from the green peptide. However, only assignment #4 has thorough product ion coverage from the blue peptide]{style="color:red"}

### Question 13

There is a reference structure for the rabbit Ribosome, PDB code:
[6hcj](https://www.rcsb.org/structure/6HCJ){target="_blank"}

The table contains the residue numbers and the PDB chain codes for
putative crosslinks. Measure the C$\alpha$ to C$\alpha$ distances of the
crosslink assignments. Note: In
[ChimeraX](https://www.cgl.ucsf.edu/chimerax/){target="_blank"}, you can load the pdb
file and use the distance command in the command line like so:

    distance /w3:91@CA /E3:224@CA

For assignment #1, residue 227 on Ribosome Protein eL6 is not resolved in
the pdb structure so use residue 224, the closest mapped residue, as I
did in the example above.

Are any of the crosslink assignments consistent with the expected
distance of the DSSO crosslinker (\< 30 Å)?

[**Answer**]{style="color:red"}

[Measuring the C$\alpha$ distances of the crosslinks on the ribosome structure gives the following values:]{style="color:red"}

```{r dist, echo=F}
pep.dist <- pep.assignments %>% select(1:4)
pep.dist$distance = c(190.238, 101.581, 164.529, 9.699)
pep.dist <- pep.dist %>%
   mutate(distance = map_chr(distance, 
                         function(x) {str_c(as.character(round(x, 2)), "Å")}))
pep.dist.flex <- flextable(pep.dist, col_keys=c(names(pep.dist)[c(-2)]))
pep.dist.flex <- compose(x = pep.dist.flex, j = "assignment", 
                    value = as_paragraph(hyperlink_text(x = assignment, url = url))
)
pep.dist.flex <- border_inner_h(pep.dist.flex, part="all", border = small_border )
pep.dist.flex <- border_inner_v(pep.dist.flex, part="all", border = small_border )
pep.dist.flex <- border_outer(pep.dist.flex, part="all", border = big_border )
pep.dist.flex %>%
   bold(part = "header") %>%
   color(part = "body", j = c("assignment"), color = "dodgerblue")

```
[Only assignment #4 is consistent with the Ribosome structure and the DSSO crosslinker]{style="color:red"}

### Question 14

What aspects of the spectral annotation were most useful in determining
whether or not the crosslink was assigned correctly?

[**Answer**]{style="color:red"}

[Clearly, #4 is the correct assignment. The two most important factors in determining that the spectrum are correcty assigned are the precursor mass accuracy and having a full set of product ions from both peptides. Many of the incorrect assignments have the DSSO diagnostic ions annotated so these are not very helpful in determining whether the mactch is correct or not]{style="color:red"}


### References:
